---
title: "final report"
author: "Charlotte Fowler"
date: "12/6/2019"
output: pdf_document
---

```{r,message=FALSE}
library(tidyverse)
library(faraway)
#library(car)
library(base)
```

```{r, message=FALSE}
lawsuit_full_prof =  read_csv("./Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    sal_avg = (sal94 + sal95)/2, 
    ln_sal_avg = log(sal_avg), 
    gender = recode(gender, "0" = "female", "1" = "male"),
    clin = recode(clin, "0" = "research", "1" = "clinical"), 
    cert = recode(cert, "0" = "not certified", "1" = "certified"), 
    rank = recode(rank, "1" = "assistant", "2" = "associate", "3" = "full professor"),
    dept = recode(dept, "1" = "biochemistry", "2" = "physiology", "3" = "genetics", "4" = "pediatrics", "5" = "medicine", "6" = "surgery") ) %>% 
  filter(rank == "full professor") %>% 
  select(-c(sal94, sal95, rank, id, sal_avg))
  



lawsuit_associate =  read_csv("./Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    sal_avg = (sal94 + sal95)/2, 
    ln_sal_avg = log(sal_avg), 
    gender = recode(gender, "0" = "female", "1" = "male"),
    clin = recode(clin, "0" = "research", "1" = "clinical"), 
    cert = recode(cert, "0" = "not certified", "1" = "certified"), 
    rank = recode(rank, "1" = "assistant", "2" = "associate", "3" = "full professor"),
    dept = recode(dept, "1" = "biochemistry", "2" = "physiology", "3" = "genetics", "4" = "pediatrics", "5" = "medicine", "6" = "surgery") ) %>% 
  filter(rank == "associate") %>% 
  select(-c(sal94, sal95, rank, id, sal_avg))


lawsuit_assistant =  read_csv("./Lawsuit.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    sal_avg = (sal94 + sal95)/2, 
    ln_sal_avg = log(sal_avg), 
    gender = recode(gender, "0" = "female", "1" = "male"),
    clin = recode(clin, "0" = "research", "1" = "clinical"), 
    cert = recode(cert, "0" = "not certified", "1" = "certified"), 
    rank = recode(rank, "1" = "assistant", "2" = "associate", "3" = "full professor"),
    dept = recode(dept, "1" = "biochemistry", "2" = "physiology", "3" = "genetics", "4" = "pediatrics", "5" = "medicine", "6" = "surgery") ) %>% 
  filter(rank == "assistant") %>% 
  select(-c(sal94, sal95, rank, id, sal_avg))
```

```{r}
lawsuit_full_prof =  read_csv("./lawsuit_full_prof.csv")
full_prof_mod_prate = lm(ln_sal_avg ~ ., lawsuit_full_prof)
vif(full_prof_mod_prate)


full_prof_mod = lm(ln_sal_avg ~ . - prate , lawsuit_full_prof)

summary(full_prof_mod)

## Model diagnosis
par(mfrow=c(2,2))
plot(full_prof_mod)

shapiro.test(residuals(full_prof_mod))
vif(full_prof_mod)

rstandard(full_prof_mod)

brand_hat =hatvalues(full_prof_mod)
brand_hat[brand_hat > 0.24]

# check influencial ponits

which(cooks.distance(full_prof_mod) > 0.5)

```

The residual has constant variance and the the residuals are normally distributed. Since the VIF is less than 5, we do not have issue with multi
collinearity. 

```{r}
lawsuit_assistant =  read_csv("./lawsuit_assistant.csv")
assistant_mod_prate = lm(ln_sal_avg ~ ., lawsuit_assistant)
vif(assistant_mod_prate)

# remove prate
assistant_mod = lm(ln_sal_avg ~ . - prate , lawsuit_assistant)



summary(assistant_mod)

## Model diagnosis
par(mfrow=c(2,2))
plot(assistant_mod)

shapiro.test(residuals(assistant_mod))

vif(assistant_mod)

rstandard(assistant_mod)

brand_hat =hatvalues(assistant_mod)
brand_hat[brand_hat > 0.18]

# check influencial ponits
which(cooks.distance(assistant_mod) > 0.5)


lawsuit_assistant1 = lawsuit_assistant [-c(68),]
assistant_mod1 = lm(ln_sal_avg ~ . - prate , lawsuit_assistant1)
summary(assistant_mod1)

```

The residual has constant variance and the the residuals are normally distributed. Since the VIF is less than 5, we do not have issue with multi
collinearity. 


Observation 68 is an influencial point, after removing this point adj r2 increase from 90% to 93%. The coefficients of department of medicine, pediatrics, gender male and clinical research change more than 10%.

```{r}
lawsuit_associate =  read_csv("./lawsuit_associate.csv")
associate_mod_prate = lm(ln_sal_avg ~ ., lawsuit_associate)
vif(associate_mod_prate)

# remove prate
associate_mod = lm(ln_sal_avg ~ . - prate , lawsuit_associate)

summary(associate_mod)

## Model diagnosis
par(mfrow=c(2,2))
plot(associate_mod)

shapiro.test(residuals(associate_mod))
vif(associate_mod)

rstandard(associate_mod)

brand_hat =hatvalues(associate_mod)
brand_hat[brand_hat > 0.3125]

# check influencial ponits
which(cooks.distance(associate_mod) > 0.5)
```

The residual has constant variance and the the residuals are normally distributed. Since the VIF is less than 5, we do not have issue with multi
collinearity. 

```{r, message=FALSE}
library(sjPlot)
library(sjmisc)
library(sjlabelled)

tab_model(full_prof_mod, associate_mod, assistant_mod1, title = "Modeling by Stratify", dv.labels = c("Ln avg_salary for Full Professor", "Ln avg_salary for Associate Professor", "Ln avg_salary for Assistant Professor" ))
```

